Describe and Explore: Using feature-based time series clustering to identify meaningful structures in intensive longitudinal data

### 2023-01-01 manuscript.tex
:sparkles: intro str bullets

% \begin{mdframed}[style=mdfbox]
% \noindent\center\textit{Structure}:
% \begin{itemize}[nosep]
%     \item introduction
%     \begin{itemize}[nosep]
%         \item\ [Relevance]: increase in number and variety of ESM data
%         \item\ [Problem]: no descriptive / understand structure analyses
%         \item\ [Proposal]: feature based clusterting = flexible +  transparent
%     \end{itemize}
%     \item Modern ESM data: new type of data and issues for analysis:
%     \begin{itemize}[nosep]
%         \item incomplete data (complete data models vs. structural missingness)
%         \item non-equidistant
%         \item non-stationary
%         \item changes/trends on different time scales
%         \item interest in multivariate developments
%         \item dimensionality curse
%     \end{itemize}
%     \item Time series clustering
%     \begin{itemize}[nosep]
%         \item shape-based
%         \item model-based
%         \item \textbf{feature-based}
%     \end{itemize}
%     \item Feature-Based Clustering
%     \begin{itemize}[nosep]
%         \item input variables
%         \item feature extraction
%         \item feature reduction
%         \item feature clustering
%     \end{itemize}
%     \item Empirical Illustration
%     \begin{itemize}[nosep]
%         \item Data
%         \item Analysis and Results
%     \end{itemize}
%     \item Discussion (summary, limitation, implications, conclusion, etc.)
% \end{itemize}
% \vspace{1em}
% \end{mdframed}

>>>>

\begin{mdframed}[style=mdfbox]
\noindent\center\textit{Structure}:
\begin{itemize}[nosep]
    \item introduction
    \begin{itemize}[nosep]
        \item increase in number and variety of ESM data + ts clustering to understand complex developmental differences.
        \item why cluster time series? 
        \begin{enumerate}[nosep]
            \item differences between clusters
            \item patterns within clusters
            \item relation to out-of cluster variables
            \item robust theories and interventions
        \end{enumerate}
        \item can't cluster all the raw data (b/c too much data and not directly comparable).
        \item summarize the data and cluster based on the summaries.
        \item ideally these summaries are:
        \begin{enumerate}[nosep]
            \item good at captuing the data
            \item meaningful = help us understand how the clustering arrived at the output (i.e., explainable) + helps us make sense of the final cluster output (i.e., interpretable)
        \end{enumerate}
        \item currently model parameters as summaries.
        \item limitations of parameter summaries: assumptions not always good at capturing data and model parameters not always interpretable
        \item current solution: more complicated models (e.g., address violated assumptions or add different types of parameters)
        \item proposed solution: features as flexible and meaningful approach of summarizing the time series
        \item ---------------------------------------------------------------------------
        \item limitations of model parameters: 
        \begin{itemize}[nosep]
            \item strong assumptions = not always a good (dis-)similiarity: 
            \begin{enumerate}[nosep]
                \item non-equidistant measurements
                \item non-imputable missingness
                \item non-stationary
                \item non-linear developments
            \end{enumerate}
            \item difficult to interpret psychologically 
            \begin{enumerate}[nosep]
                \item limited to few ways of psychological effects (e.g., lagged + cross-lagged effects)
            \end{enumerate}
        \end{itemize} 
        \item push for more complicated models aimed at addressing (1) similarity concerns (e.g., amend assumptions; e.g., see den Teuling) or (2) interpretatbility (e.g., more types of effects; e.g., see krone2018)
        \item proposed alternative: go back to interpretable similarity measurements. General statistics and parameters instead of only parameters.
    \end{itemize}
    \item Feature-based Time Series Clustering
    \begin{itemize}[nosep]
        \item different types of similarity grouping: five types
        \item why features: flexible similarity (b/c more variety), less assumptions, easy to calculate/extract, theory-based selection (as per RQ), psychological interpretability, ...
        \item Features in Psychological Time Series
        \begin{itemize}[nosep]
            \item central tendency
            \item variability
            \item (in)stability
            \item self-similarity
            \item linear trend
            \item nonlinearity
        \end{itemize}
        \item Analysis Steps
        \begin{itemize}[nosep]
            \item input variables
            \item feature extraction
            \item (feature reduction)
            \item feature clustering
            \item cluster evaluation
        \end{itemize}
    \end{itemize}
    \item Empirical Illustration
    \item Discussion (summary, limitation, implications, conclusion, etc.)
\end{itemize}
\vspace{1em}
\end{mdframed}
